---
title: "glmertree updates"
author: "Marjolein Fokkema"
date: "17 april 2019"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Convergence criterion and contrast coding

The convergence check assumed that the log-likelihood value could only increase from one iteration to the next. In the earlier version of **glmertree**, if the log-likelihood decreased (which is quite likely if the tree size decreases from one iteration to the next), estimation converged even though the log-likelihood values are quite different:

```{r, warning=FALSE, message=FALSE}
library("glmertree", lib.loc = "./glmertree 0.1-2/")
```

```{r, warning=FALSE, message=FALSE}
lt <- lmertree(depression ~ treatment | cluster | age + anxiety + duration,
               data = DepressionDemo, verbose = TRUE)
```

This yields differences between the estimated coefficients from `lmtree` and `lmer`:

```{r}
coef(lt$tree)
fixef(lt$lmer)
```

Also, the contrasts used by `lmtree` and `lmer` differ. `lmer` uses treatment coding, where the first level serves as a reference category for the intercepts in the other nodes, whereas `lmtree` estimates a separate intercept for each node.

The convergence and coding issues are corrected in the current development version:

```{r, warning=FALSE, message=FALSE}
detach("package:glmertree", unload=TRUE)
library("glmertree", lib.loc = "./glmertree 0.1-3/")
```

For checking convergence, the absolute difference in log-likelihood values from one iteration to the next is checked in `(g)lmertree`:

```{r}
lt <- lmertree(depression ~ treatment | cluster | age + anxiety + duration,
               data = DepressionDemo, verbose = TRUE)
```

We now have one additional iteration, which yields identical estimates of `lmtree` and `lmer`:

```{r}
coef(lt$tree)
fixef(lt$lmer)
```

Also, `lmtree` and `lmer` now use identical coding. 

Also, the `coef` and `fixef` methods of `lmertree` now return the `lmer` estimates:

```{r}
fixef(lt)
coef(lt)
```

These changes have also been implemented for `glmertree`.


# Plotting

`plot.(g)lmertree` plots the fitted values, which are also based on the (varying) random effects predictions, as well as the effects of additional predictors of the linear model. This yields wiggly lines for continuous predictors, e.g.:

```{r}
lt4 <- lmertree(depression ~ treatment + age | cluster | anxiety + duration,
               data = DepressionDemo)
plot(lt4, which = "tree")
```

The wiggly lines may be somewhat counter-intuitive, especially when partitioning growth curve models. E.g., an example based on an artificial example dataset for growth curve modeling I included in the package:

```{r}
form <- formula(paste0("y ~ time | person | ", 
                paste0("x", 1:28, collapse = " + ")))
lt.growth <- lmertree(form, cluster = person, data = GrowthCurveDemo)
plot(lt.growth, which = "tree")
```

I wrote a new plotting function which uses an adjusted version of `node_bivplot` from **partykit**. For plotting the fitted values, it allows for assuming the random effects to be constant (at the population mean of 0) and the fixed effects to be constant (at their sample means), through the following arguments:

* `fit.ranef`: character; `"fixed"` (default) or `"varying"`. Only used when `fitmean = TRUE`. If `"constant"`, fitted means will be calculated, with random effects fixed at the population mean of 0. If `"varying"`, fitted means will also be based on the fitted random effects.
* `fit.fixef`: character, `"constant"` (default) or `"varying"`. Only used when `fitmean = TRUE`. If `"constant"`, fitted means will be calculated, keeping the values of the other predictors in the (G)LM constant at their sample mean or majority value. If `"varying"`, fitted means will also be based on the fitted effects of the remaining predictor variables.

The new function plots straight lines, by default:

```{r}
plot.lmertree2(lt.growth, which = "tree")
```

The earlier default behavior (fitted values include variation due to random and fixed effects) can be requested as follows:

```{r}
plot.lmertree2(lt.growth, which = "tree", fit.ranef = "varying", fit.fixef = "varying")
```

We can also omit the fitted values:

```{r}
plot.lmertree2(lt.growth, which = "tree", fitmean = FALSE)
```

Or omit the observed values:

```{r}
plot.lmertree2(lt.growth, which = "tree", observed = FALSE)
```

Or only plot the coefficients:

```{r}
plot.lmertree2(lt.growth, which = "tree", type = "simple")
```

By specifying `which = "both"` or `which = "ranef"`, the (co)variances of the random effects would (also) be printed in each of the terminal nodes.

Function `plot.glmertree2` provides similar behavior for `glmertree`s. The new function has separate documentation to explain the arguments, see:

```{r, eval=FALSE}
?plot.lmertree2
```


# Updated vignette

The original vignette only included an example on treatment-subgroup interactions. I get quite some emails from users who do not know how to specify other types of models for the nodes, so I have extended the vignette with two additional examples: one with constant fits in the terminal nodes and one with growth curve models in the terminal nodes. see: 

```{r}
vignette("glmertree", package = "glmertree")
```

# Unresolved: Differences between fixef(glmer) and coef(glmtree)

Above, we saw that `fixef(lmer)` and `coef(lmtree)` yield (nearly) identical results in the last iteration of `lmertree`:

```{r}
coef(lt) - coef(lt$tree)
```

But this is not the case for `glmertree`, `glmtree` and `glmertree`:

```{r}
gt <- glmertree(depression_bin ~ treatment | cluster | age + anxiety + duration,
                data = DepressionDemo)
coef(gt) - coef(gt$tree)
```

My current best guess is that the approach of `(g)lmertree`, i.e., estimating the tree with the random effects predictions as an offset, is more similar to REML than to ML estimation of a mixed-effects model. And thus, the estimated coefficients of `lmertree` are identical to those of `lmer` with REML estimation (the default):

```{r}
## Estimate coefficients as lmtree does in lmertree:
lm1 <- lm(depression ~ .tree + .tree:treatment - 1 + offset(.ranef), 
          data = lt$data) 

## Estimate coefficients as lmer does in lmertree:
lmm.REML <- lmer(depression ~ .tree + .tree:treatment - 1 + (1|cluster), 
                 data = lt$data)
coef(lm1) - fixef(lmm.REML) ## (nearly) identical 
```

but not to those of `lmer` with ML estimation:

```{r}
lmm.ML <- lmer(depression ~ .tree + .tree:treatment - 1 + (1|cluster), 
               data = lt$data, REML = FALSE)
coef(lm1) - fixef(lmm.ML) ## somewhat different 
```

The larger differences between `glmer` and `glmtree` may then occur, because estimation of GLMMs is more complex altogether, and `glmer` does not support REML, only ML estimation. According to http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#reml-for-glmms, it is unclear how to define the REML criterion in GLMMs. Perhaps because there are different ways to define the residuals with a binary respose.

# Vignette extended

The original vignette contains an examples on detecting treatment-subgroup interactions only. I regularly get e-mail from users who do not know how to specify other models. So I added examples on fitting GLMM trees with constant fits in the terminal nodes (section 2) and partitioning of growth-curve models (section 4):

```{r, eval=FALSE}
vignette("glmertree", package = "glmertree")
```