---
title: "glmertree updates"
date: "15 May 2019"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("glmertree")
```

## Global and local fixed effects

`coef` and `fixef` methods now have an additional argument: `which`:
 
```{r}
## fit LMM tree including both local and global fixed effect
lt <- lmertree(depression ~ treatment | (age + (1|cluster)) | anxiety + duration,
  data = DepressionDemo)
coef(lt, which = "tree") # default
coef(lt, which = "global")
```

Also, they have an argument `drop`, which is `FALSE`, by default. It was somewhat easier to always have a matrix of coefficients returned, with a row for every terminal node and a column for every local fixed effect.

Separate documentation for these methods has now been included and can be accessed through any of the following commands:

```{r, eval=FALSE}
?coef.lmertree
?coef.glmertree
?fixef.lmertree
?fixef.glmertree
``` 




## Plotting coefficients without tree structure

### Default settings

```{r}
lt <- lmertree(depression ~ treatment + age | cluster | anxiety + duration,
  data = DepressionDemo)
plot.lmertree2(lt, which = "tree.coef")
plot.lmertree2(lt, which = "ranef")
```


### No error bars when joint = FALSE

```{r}
lt <- lmertree(depression ~ treatment + age | cluster | anxiety + duration,
  data = DepressionDemo, joint = FALSE)
plot.lmertree2(lt, which = "tree.coef")
```


### Warning when global fixef are specified

```{r}
lt <- lmertree(depression ~ treatment | (age + (1|cluster)) | anxiety + duration,
  data = DepressionDemo)
plot.lmertree2(lt, which = "tree.coef")
```