---
title: "glmertree updates"
date: "7 May 2019"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## New convergence checks

```{r}
library("glmertree")
load(file = "ex_data.Rda")
lt_form1 <- HAMD ~ time * Condition | (1|Study/PatientID) | Gender + 
  Age + Zanx + rawHAMDpre
lmt1 <- lmertree(lt_form1, parm = 5:6, verbose = TRUE, 
                 data = ex_data)
lmt2 <- lmertree(lt_form1, parm = 5:6, verbose = TRUE, maxdepth = 3L, 
                 data = ex_data)
```

Note that in the first tree, estimation will start to go round in circles, and therefore is stopped after the third iteration. Perhaps we should print a warning, as the estimated coefficients from the `lmtree` and `lmer` will differ (see below).

In the second tree, estimation converged normally.


## Effects coding and fixef, coef methods

I have undone the earlier changes to the coding `lmer` uses for the fixed effects. The `fixef` and `coef` methods for `(g)lmertree` are a bit tweaky now, but return the `(g)lmer` estimates, in the shape of the `(g)lmtree` `coef` method:

```{r}
coef(lmt1)
coef(lmt1$tree)
fixef(lmt1$lmer)
```

Note that estimation did not converge for the first tree, so the estimates of `lmtree` and `lmer` differ somewhat. If estimation did converge, all estimated parameters are the same, and differ only because of the different effects coding schemes:

```{r}
coef(lmt2)
coef(lmt2$tree)
fixef(lmt2$lmer)
```


## Plotting

Current default:

```{r}
plot(lmt1, which = "tree", gp = gpar(cex = .5))
```

We could include option `which = "coef"`, in which case coefficients from the tree and random effects would be plotted. For tree coefficients, we can use `dotchart()` from base **`graphics`**, or `dotplot()` from **`lattice`**: 

```{r}
coefs <- coef(lmt1)
dotchart(coefs, labels = paste("node", coefs),
         xlab = "Estimated coefficients",
         main = "Fixef effects from tree")
long_coefs <- data.frame(stack(data.frame(coefs)), 
                         node = rep(paste("node", rownames(coefs)), 
                                    times = ncol(coefs)))
lattice::dotplot(node ~ values | ind, data = long_coefs, as.table=TRUE, 
                 xlab = "Estimated coefficients",
                 main = "Fixed effects from tree")
```

The latter may be preferred, as it provides the same plot style as for the random effects:   

```{r}
plot(lmt1, which = "ranef")
```

Although `dotplot()` introduces a dependency on **`lattice`**, this is already imported by **`lme4`** anyway.





```{r}
lt.growth <- lmertree(y ~ time | person | x1 + x2 + x3 + x4 + x5, 
                      cluster = person, data = GrowthCurveDemo)
plot.lmertree2(lt.growth, type = "simple", terminal_panel = node_terminal_glmertree)
```


```{r}
plot.lmertree2(lt.growth, fitmean = "none")
```

```{r}
plot.lmertree2(lt.growth, fitmean = "combined")
```

```{r}
plot.lmertree2(lt.growth, fitmean = "marginal")
```